---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/"
)
```

# puddingR

[![lifecycle](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
 
`puddingR` is designed to streamline R data analysis and open data release for the team at The Pudding. Check out the full `pkgdown`-generated site [here](https://link).
 
Learn more by searching `vignette("puddingR")` in your R or RStudio console.
 
## Installation

At the moment, there are no plans for `puddingR` to appear on CRAN. To download:

```{r eval = FALSE}
# Install development version from GitHub
devtools::install_github("the-pudding/puddingR")

# Load into your R environment
library(puddingR)
```

## Usage

The `puddingR` package currently has a few main features:

### Create a new project & directory system

For our front-end work, the team at The Pudding, uses a [starter template](https://github.com/the-pudding/starter) which sets up anything that could be useful for our work and creates a file directory system. The directory helps the team (and the author) know exactly where to find everything they need for a project. 

The `create_project()` function does the same thing for R projects. It creates a single folder, by default called `analysis`, creates a new R project and several directories. The resulting structure looks like this: 

```{r results="asis", eval = FALSE}
analysis
|-- assets
  |-- data
    |-- open_data
    |-- processed_data
    |-- raw_data
|-- packrat
|-- plots
|-- R
|-- reports
|-- rmds
|-- analysis.Rproj
```

Many of the remaining functions in this package use default values that assume the above folder structure is present. 

```{r eval = FALSE}
puddingR::create_project(title = "My Cool Project")
```


### Create a Pudding-themed Rmd Report

While RStudio comes with several themes to use for Rmarkdown reports, I have created one specifically for use with Pudding projects. It contains our logo and lots of tips for how to import and display data. 

To use it, in RStudio, navigate to
`File` > `New File` > `Rmarkdown`

A window will pop up with a list of options on the left-hand side. Select `From Template` from that list, and navigate to `Pudding Report`. This will launch open a new Rmd file filled with tips that can be deleted if you don't need them. If you're using the above file structure, save this into the `rmds` folder. 

### Outputting Data

Often in our work we need to write data that we have analyzed and save it in another location. In this workflow, most processed data belongs in `assets/data/processed_data/`. The easiest way to get it there is to use `export_data()`. This function has several arguments, included in them is `location`. If you are using the default file structure from `puddingR`, you can simply use `open` to save it to the `open_data` directory, `processed` to save it to the `processed_data` directory, or, if this `analysis` folder exists in the same working directory as the front-end Pudding starter template, you can use `js` (which will save the file in `src/assets/data/` for use in front-end work).

```{r eval = FALSE}
# To export mtcars dataset to the processed_data directory
puddingR::export_data(mtcars, "cars_data", location = "processed", directory = "auto", format = "csv")
```


### Outputting Data + Metadata

If you want to export codebooks and the scripts used to generate data at the same time the data is generated, you can use `export_all()` instead. 

```{r eval = FALSE}
# To export mtcars dataset to the open_data directory
puddingR::export_all(mtcars, "cars_data", 
                      location = "open", 
                      directory = "auto", 
                      format = "csv", 
                      # Also export a codebook to accompany the data  
                      codebook = TRUE,
                      # Export the code chunks from an Rmd file used to generate the data
                      scripts = c("load_data", "analyze_data", "graph_data"),
                      scriptFile = here::here("rmds/analysis.Rmd")
                      )
```

These pieces can be accomplished separately using `export_data()`, `create_codebook()` and `export_code()`.

The default directory for a codebook is in `open_data/intermediates`. These files should be opened and manually edited to provide the correct metadata to accompany an open data file. 

Once all of the intermediate codebooks have been edited, they can all be combined into a single file using `knit_data_readme()`.

```{r eval = FALSE}
puddingR::knit_data_readme()
```

Find more information on how to use this package in the vignettes or on the `pkgdown` site. 
 
## Code of Conduct

Please note that this project is released with a [Contributor Code of Conduct](CODE_OF_CONDUCT.md).
By participating in this project you agree to abide by its terms.
